Cluster Difference Imaging Photometric Survey. Vetting Report Description Document

### 2020-03-04 doc/vetting_report_documentation/doc_TeX/ms.tex
update vetting report description document
I now understand why changing anything in these vetting reports is a big
hassle for NASA. You need to rewrite the documentation too! And it's
never backwards-compatible!

\begin{abstract}
To find planet candidates in clusters,
we make vetting reports using our light-curves \citep{bouma_cdips_2019} and
auxiliary data.
This document describes the CDIPS planet candidate vetting reports
uploaded by \texttt{lbouma} to ExoFOP-TESS 2019-12-15 through present.
Earlier versions of this document are available at
  \url{lgbouma.com}.
\\

@@

Figures~\ref{fig:pg1} to~\ref{fig:pg7} summarize the document
construed for these purposes.  The planet candidate chosen for these
figures (Gaia-DR2 \texttt{532542547468722176} = TIC
\texttt{401153825}) was chosen because it showcases many of the
vetting report features, while also being a likely eclipsing binary.

@@

Figure~\ref{fig:pg2}.
Time-series of raw flux (\texttt{IRM2}),  TFA-detrended flux
(\texttt{TF2}), stellar-variability detrended flux, and the background
are shown as a function of barycentric Julian date.  The overplotted
dashed vertical lines are the ephemeris of the highest-power TLS peak
from Figure~\ref{fig:pg1}.  An important visual check is whether the flux
dips are correlated with changes in the background level -- in this
case, they are not.  The standard deviation and TESS magnitude are
quoted in the upper right.  The red line in the second from the top
plot is a spline fit, which in this case was not an essential step for
finding the eclipse signal.

The spline is an optional feature: it is only fitted and removed if
the star is found to be ``variable'' (the Lomb-Scargle peak period is
found with false alarm probability below $10^{-5}$). The spline is a
robust penalized B-spline, which is a B-spline with knot-length
automatically determined via cross-validation
\citep{eilers_flexible_1996}. The idea behind the cross-validation is
that more knots leads to smaller residuals on training data, but
larger errors when tested on the entire dataset.  We used the
\texttt{wotan} implementation, which is a wrapper to the
\texttt{pyGAM} spline fitter, with $2\sigma$ clipping of outliers from
the fit residuals at each iteration
\citep{serven_pygam_2018_1476122,hippke_wotan_2019}.  The maximum
number of spline knots was set to 50, which for each TESS sector
($\approx 25\,{\rm days}$) is commensurate with a $\approx0.5\,{\rm
day}$ window.

@@

Figure~\ref{fig:pg4}.
Apertures of radius 1, 1.5, and 2.25 pixels are shown from top to
bottom.  The blue line is the reference transit depth from the
best-fitting TLS model.  Changes in depth with increasing aperture
size can indicate that the source of variability is off-center from
the aperture, suggesting a photometric blend.

@@

Figure~\ref{fig:pg5}.
The star was considered a candidate cluster member by the source(s)
listed under ``Reference'', in this case
\citet{cantat-gaudin_gaia_2018}.  The name used in their catalog was
\texttt{5325425474687221760}, a Gaia-DR2 identifier, which can be
back-referenced to find the membership probability they assigned this
star for being in Collinder 205.  The base catalog however for this page
of plots is \citet{Kharchenko_et_al_2013}, due to its homogeneous
parameter determination procedure (particularly for age).  If a match to
the \citet{Kharchenko_et_al_2013} catalog is found, then the subplots on
this page are populated; otherwise, they are not.  Top-left shows the
parallax, with orange points sampled from the Gaia-DR2 posterior, black
points the other cluster members in the {\it Kharchenko catalog} (not
the catalog claiming membership), and the blue line the claimed
Kharchenko parallax for the cluster.  A number of field contaminants in
the Kharchenko catalog are visible in this case.  Top-right are the Gaia
proper motions, where against black points are cluster members from
Kharchenko, and the orange is the target star.  Bottom-left is the
color-magnitude diagram, and bottom-right are the on-sky positions.  In
the text, \texttt{N1sr2} is the number of $1\sigma$ cluster members
reported by \citet{Kharchenko_et_al_2013} within the cluster angular
radius; $\log t$ is the base-10 logarithm of the age in years;
\texttt{type} matches the type codes provided by
\citet{Kharchenko_et_al_2013}; ``\texttt{Note}'' gives the description
of the cluster from \citet{Kharchenko_et_al_2013}, if any is available.
Extra caution must be taken when interpreting this set of plots, since
they can only show disagreement between the observed star's properties
and those of the listed Kharchenko members (and the latter may be
incomplete or otherwise biased).

@@

Figure~\ref{fig:pg7} shows the labeled quantities from the target
star, the neighborhood, and the ``cluster members'' reported by
\citet{cantat-gaudin_gaia_2018}, \cite{kounkel_untangling_2019}, or
\citet{Kharchenko_et_al_2013}.  The top three subplots intentionally
omit the labelled cluster members, in order to give the user their own
by-eye assessment of whether they see clusters in the neighborhood (as
might happen if the membership labelling is {\it incomplete}),
and whether the target star is within those clusters.
The middle three subplots overplot cluster members from the reference
noted in the legend.

>>>>

\begin{abstract}
To find planet candidates in clusters, we make vetting reports using
our light-curves \citep{bouma_cdips_2019} and auxiliary data.  This
document describes the CDIPS planet candidate vetting reports
uploaded by \texttt{bouma} to ExoFOP-TESS 2020-01-20 through
2020-05-01.  Earlier versions of this document are available at
\url{lgbouma.com/notes}.
\\

@@

Figures~\ref{fig:pg1} to~\ref{fig:pg7} summarize the document
construed for these purposes.  The planet candidate chosen for these
figures (Gaia-DR2 \texttt{3050033749239975552} = TIC
\texttt{125192758}) was chosen because it showcases many of the
vetting report features, in this case for a neighboring eclipsing binary.

@@

Figure~\ref{fig:pg2}.
Time-series of raw flux (\texttt{IRM2}),  TFA-detrended flux
(\texttt{TF2}), stellar-variability ``detrended flux'', and the background
are shown as a function of barycentric Julian date.  The overplotted
dashed vertical lines are the ephemeris of the highest-power TLS peak
from Figure~\ref{fig:pg1}.  An important visual check is whether the flux
dips are correlated with changes in the background level -- in this
case, they are not.  The standard deviation and TESS magnitude are
quoted in the upper right.  The red line in the top
plot is a spline fit, which in this case helped in
finding the eclipse signal.

The spline is an optional feature: it is only fitted and removed if
the star is found to be ``variable'' (the Lomb-Scargle peak period is
found with false alarm probability below $10^{-5}$, within the TFA light-curves). The spline is a
robust penalized B-spline, which is a B-spline with knot-length
automatically determined via cross-validation
\citep{eilers_flexible_1996}. The idea behind the cross-validation is
that more knots leads to smaller residuals on training data, but
larger errors when tested on the entire dataset.  We used the
\texttt{wotan} implementation\footnote{
We use \texttt{wotan} versions greater than \texttt{1.4}. Before \texttt{2020-03-01}, an earlier version
was used, which led to qualitatively worse detrending performance for a subset
of light-curves due to a floating point truncation error. See \texttt{\url{https://github.com/hippke/wotan/issues/51}}.
}, which is a wrapper to the
\texttt{pyGAM} spline fitter, with $2\sigma$ clipping of outliers from
the fit residuals at each iteration
\citep{serven_pygam_2018_1476122,hippke_wotan_2019}.  The maximum
number of spline knots was set to 50, which for each TESS sector
($\approx 25\,{\rm days}$) is commensurate with a $\approx0.5\,{\rm
day}$ window.

Whenever the TFA light-curve is found to be variable, the ``stellar variability detrended flux'' is
the result of dividing the raw \texttt{IRM2} flux by the best-fit spline\footnote{
Before \texttt{2020-03-01}, this definition was different. For the initial
single-sector processing of S6-S11, it was the result of dividing the \texttt{TFA2} flux
by the best-fit spline.
}.
This is the light-curve that is then searched for transiting planets, and used
in {\it e.g.}, Figure~\ref{fig:pg1} and Figure~\ref{fig:pg3}.
If the TFA light-curve is not found to be variable, then the \texttt{TFA2} light-curve is used for
this purpose.

@@

Figure~\ref{fig:pg4}.
Apertures of radius 1, 1.5, and 2.25 pixels are shown in maximally detrended
light-curves from top to
bottom.  The blue line is the reference transit depth from the
best-fitting TLS model.  Changes in depth with increasing aperture
size can indicate that the source of variability is off-center from
the aperture, suggesting a photometric blend.

@@

Figure~\ref{fig:pg5}.
The star was considered a candidate cluster member by the source(s)
listed under ``Reference'', in this case
\citet{cantat-gaudin_gaia_2018}.  The name used in their catalog was
in this case a Gaia-DR2 identifier, which can be
back-referenced to find the membership probability they assigned this
star for being in NGC 2323.  The base catalog however for this page
of plots is \citet{Kharchenko_et_al_2013}, due to its homogeneous
parameter determination procedure (particularly for age).  If a match to
the \citet{Kharchenko_et_al_2013} catalog is found, then the subplots on
this page are populated; otherwise, they are not.  Top-left shows the
parallax, with orange points sampled from the Gaia-DR2 posterior, black
points the other cluster members in the {\it Kharchenko catalog} (not
the catalog claiming membership), and the blue line the claimed
Kharchenko parallax for the cluster.  A number of field contaminants in
the Kharchenko catalog are visible in this case.  Top-right are the Gaia
proper motions, where against black points are cluster members from
Kharchenko, and the orange is the target star.  Bottom-left is the
color-magnitude diagram, and bottom-right are the on-sky positions.  In
the text, \texttt{N1sr2} is the number of $1\sigma$ cluster members
reported by \citet{Kharchenko_et_al_2013} within the cluster angular
radius; $\log t$ is the base-10 logarithm of the age in years;
\texttt{type} matches the type codes provided by
\citet{Kharchenko_et_al_2013}; ``\texttt{Note}'' gives the description
of the cluster from \citet{Kharchenko_et_al_2013}, if any is available.
Extra caution must be taken when interpreting this set of plots, since
they can only show disagreement between the observed star's properties
and those of the listed Kharchenko members (and the latter is in some cases
incomplete or otherwise biased).

@@

Figure~\ref{fig:pg7} shows the labeled quantities from the target
star, the neighborhood, and the ``cluster members'' reported by
\citet{cantat-gaudin_gaia_2018}, \cite{kounkel_untangling_2019}, or
\citet{Kharchenko_et_al_2013}.  The top three subplots intentionally
omit the labelled cluster members, in order to give the user their own
by-eye assessment of whether they see clusters in the neighborhood (as
might happen if the membership labelling is {\it incomplete}),
and whether the target star is within those clusters.
The middle three subplots overplot cluster members from the reference
noted in the legend.
The radial velocity plots (bottom row) use radial velocities from Gaia-DR2,
and so are limited to roughly 3500-7000$\,$K, for $G\lesssim 12.5$ \citep{katz_gaia_2019}.
